---
title: "Little 500 Project"
author: "Jayden Fisher"
date: "2023-11-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading Libraries and Data.
```{r}
library(dplyr)
library(tidyr)
library(lubridate)

itt_data = read.csv("Individual_Time_Trials.csv")
little500_data = read.csv("Little_500_Race.csv")
qualifications_data = read.csv("Qualifications.csv")
team_pursuit_data = read.csv("Team_Pursuit.csv")
```

Changing Column Names, and only using Male race's for consistency (Female times are much different since only 100 laps.)
```{r}
itt_data = itt_data %>%
  filter(Gender == 'M') %>%
  rename(Team = Team, Finish_Time = `Finish.Time`)

little500_data = little500_data %>%
  filter(Gender == 'M') %>%
  rename(Team = Team, Finish_Time_Little500 = `Finish.Time`, Laps = Laps)

qualifications_data = qualifications_data %>%
  filter(Gender == 'M') %>%
  rename(Team = Team, Finish_Time_Qualifications = `Finish.Time`)

team_pursuit_data = team_pursuit_data %>%
  filter(Gender == 'M') %>%
  rename(Team = Team, Finish_Time_Team_Pursuit = `Finish.Time`)

model_data = little500_data
```

Cleaning team ITT, getting averages between years, converting to seconds, adding to model data.
```{r}
itt_data = itt_data %>%
  mutate(
    Finish_Time = as.numeric(ms(Finish_Time))  
  )

itt_data = itt_data %>%
  group_by(Team, Year) %>%
  summarize(Avg_ITT_Time = mean(Finish_Time, na.rm = TRUE)) %>%
  ungroup()

itt_data = na.omit(itt_data)

model_data = merge(model_data, itt_data, by = c("Team", "Year"), all.x = TRUE)
```

Cleaning team Qualifying, getting averages between years, converting to seconds, adding to model data.
```{r}
qualifications_data = qualifications_data %>%
  mutate(
    Finish_Time_Qualifications = as.numeric(ms(Finish_Time_Qualifications))  
  )

qualifications_data = qualifications_data %>%
  group_by(Team, Year) %>%
  summarize(Finish_Time_Qualifications = mean(Finish_Time_Qualifications, na.rm = TRUE)) %>%
  ungroup()

qualifications_data = na.omit(qualifications_data)
model_data = merge(model_data, qualifications_data, by = c("Team", "Year"), all.x = TRUE)
```

Cleaning team pursuit, getting averages between years, converting to seconds, adding to model data.
```{r}
team_pursuit_data = team_pursuit_data %>%
  mutate(
    Finish_Time_Team_Pursuit = as.numeric(ms(Finish_Time_Team_Pursuit))
  )

team_pursuit_data = na.omit(team_pursuit_data)

team_pursuit_data = team_pursuit_data %>%
  group_by(Team, Year) %>%
  summarize(Finish_Time_Team_Pursuit = mean(Finish_Time_Team_Pursuit, na.rm = TRUE)) %>%
  ungroup()

model_data = merge(model_data, team_pursuit_data, by = c("Team", "Year"), all.x = TRUE)
```

Cleaning model data and converting little 500 into seconds.
```{r}
model_data = model_data %>%
  mutate(
    Finish_Time_Little500 = as.numeric(ms(Finish_Time_Little500)) 
  )

model_data = model_data %>%
  filter(!is.na(Finish_Time_Qualifications) & !is.na(Avg_ITT_Time) & !is.na(Finish_Time_Team_Pursuit) & !is.na(Finish_Time_Little500))
```

Interaction Tests.
```{r}
full_model_interaction = lm(Finish_Time_Little500 ~ Finish_Time_Qualifications * Avg_ITT_Time + Finish_Time_Team_Pursuit, data = model_data)

reduced_model_no_interaction = lm(Finish_Time_Little500 ~ Finish_Time_Qualifications + Finish_Time_Team_Pursuit + Avg_ITT_Time, data = model_data)

anova_result = anova(reduced_model_no_interaction, full_model_interaction)

anova_result
```

Model Predictor Tests.
```{r}
anova_result_1 = anova(lm(Finish_Time_Little500 ~ Avg_ITT_Time + Finish_Time_Team_Pursuit, data = model_data), full_model_interaction)

anova_result_2 = anova(lm(Finish_Time_Little500 ~ Finish_Time_Qualifications + Finish_Time_Team_Pursuit, data = model_data), full_model_interaction)

anova_result_3 = anova(lm(Finish_Time_Little500 ~ Finish_Time_Qualifications + Avg_ITT_Time, data = model_data), full_model_interaction)

anova_result_1

anova_result_2

anova_result_3
```

Final Model
```{r}
model = lm(Finish_Time_Little500 ~ Finish_Time_Qualifications * Avg_ITT_Time + Finish_Time_Team_Pursuit, data = model_data)

summary(model)
```

Testing the model
```{r}
new_data = data.frame(
  Finish_Time_Qualifications = c(150.12),
  Avg_ITT_Time = c(146.52),
  Finish_Time_Team_Pursuit = c(588.35)
)

predicted_finish_times = predict(model, newdata = new_data)

predicted_finish_times_hours = floor(predicted_finish_times / 60)
predicted_finish_times_minutes = round(predicted_finish_times %% 60, 2)

predicted_finish_times_formatted = sprintf("%02d:%05.2f", predicted_finish_times_hours, predicted_finish_times_minutes)

predicted_finish_times_formatted
```


